---
title: "  "
output: html_notebook
---

### Where does Ghana's imported fruit come from?

#### Analyzing data from Food and Agriculture Organization (FAO) bilateral trade database 1998-2019

#### By: Ama Owusu-Darko

```{r load global packages, include=FALSE}
library(tidylog)
library(tidyverse)

library(tinytex)

library(labelled)
```

```{r load data, include=FALSE}
dat<- read.csv("~/r_and_git/Food_trade/good_fruits.csv", stringsAsFactors=TRUE)



```

```{r glance data, include=FALSE}
#summary(dat)
```

```{r eda ,include=FALSE, eval=F}
dat<-as_tibble(dat)
dat$
dat1<-dat%>%
  mutate(across(c(Year,Item.Code,Element.Code,Partner.Country.Code,Reporter.Country.Code,ï..Domain.Code),as_factor))
         
#removing duplicate columns
dat2<-dat1%>%
  select(-c(Year.Code,Item.Code,Element.Code,Partner.Country.Code,Reporter.Country.Code,ï..Domain.Code))
```

```{r group_by, include=FALSE}
dat3<-dat2%>%
  filter(Element=="Import Quantity")%>%
  group_by(Partner.Countries,Item, Element, Year)
head(dat3)

#Select import quantity because the value of dollars change because of inflation over the years
dat4<-dat2%>%
  filter(Element=="Import Quantity")%>%
  group_by(Year,Partner.Countries) %>% 
  summarise(Value)
head(dat4)

```

**Assumption:**
Import quantities (measured in tonnes) are considered in the analysis



```{r countries did Ghana import from, echo=FALSE, message=FALSE, warning=FALSE}
#remove value==0
imp<-dat2%>%
  filter(Element=="Import Quantity" & Value!=0 & Unit=="tonnes")
  

im<-dat2%>%
  filter(Element=="Import Quantity" & Value!=0 & Unit=="tonnes")%>%
  count(Year)%>%
  ggplot()+geom_bar(aes(x=year))+coord_flip()


imp1<-dat2%>%
  filter(Element=="Import Quantity"& Value!=0 & Unit=="tonnes")%>%
  group_by(Partner.Countries)%>%
  summarize(Per_country=n())



```

<br> **Question: Which countries exported the most fruit to Ghana during 1998-2019?**

```{r 5 exporters, echo=FALSE, message=FALSE, warning=F}
try_c=function(fun1){
 imp1%>%
  fun1(Per_country,n=5)%>%
  ggplot(aes(x=Partner.Countries, y=Per_country, fill=Partner.Countries))+geom_col()+
  scale_fill_brewer(palette="Set3")+
    geom_label(aes(label=Per_country),color="white",fill="black")+
    labs(title="The 5 Most Frequent Exporters of fruits to Ghana",
         subtitle="Values represent count of exports to Ghana")
}

try_c(slice_max)


try_v=function(){
imp1%>%
  slice_max(Per_Country,n=5)%>%
  ggplot(aes(x=Partner.Countries, y=Per_country, fill=Partner.Countries))+geom_col()+
  scale_fill_brewer(palette="Set3")+
    geom_label(aes(label=Per_Country),color="white",fill="black")+
    labs(title="The 5 Most Frequent Exporters of fruits to Ghana",
         subtitle="Values frquency of exports to Ghana")
}

#try_v(slice_max)

#five_exp=imp1%>%
  #slice_max(Per_country,n=5)

only_top5=imp%>%
  add_count(Partner.Countries)%>%
  filter(dense_rank(desc(n))<=5)%>%
  group_by(Partner.Countries,Item)%>%
  summarize(Per_item=n())%>%
  slice_max(Per_item, n=5)
```

```{rk, include=F}
only_top5%>%
  ggplot(aes(x=Item,y=Per_item, fill=Partner.Countries))+geom_col()+
scale_fill_brewer(palette="Set3")+facet_wrap(~Partner.Countries, scales = "free", ncol=5)+
geom_label(aes(label=Per_item),color="white",fill="black")+
labs(title="The 5 Most Imported fruits from  to Ghana",
subtitle="Values represent Frequency")+theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



```{r include=FALSE}
b=nly_top5%>%
  group_by(Year,Item)%>%
  summarize(Per_item=n())%>%
  slice_max(Per_item, n=5)
```

```{r include=FALSE}
library(gtsummary)
theme_gtsummary_journal(journal = "jama")
#> Setting `JAMA` theme
theme_gtsummary_compact()
#> Setting `Compact` theme
gt1 <- 
  dat2%>%
  select(Value)%>%
  tbl_summary()
    
gt1


dat5<-dat2%>%
mutate(cat_Value = ifelse(Value>10000,"High tonnage", ifelse(Value>1000,"Medium_tonnage",ifelse(Value>100,"Small_tonnage",                                "Very_Small_tonnage"))))
  
dat5$cat_Value<-factor(dat5$cat_Value, levels=c("Very_Small_tonnage",
               "Small_tonnage","Medium_tonnage","High tonnage"))

nly_top5=imp%>%
  select(2,3,5,6,Value,Unit)%>%
  group_by(Partner.Countries,Item, Value)%>%
  filter(Value>300)

f=nly_top5%>% 
  group_by(Partner.Countries)%>%
  count(Partner.Countries)



  #filter(dense_rank(desc(n))<=5)%>%
  #group_by(Partner.Countries,Item)%>%
 # summarize(Per_item=n())%>%
 # slice_max(Per_item, n=5)
#d=imp%>%
  
 # tidycomm::crosstab(Year,Partner.Countries,Item,Value,add_total = TRUE)
```

**Question: How many tons of fruit does Ghana normally import?**


Records of nine fruits were selected from Food and Agriculture Organization (FAO) bilateral trade database 1998-2019.

```{r 1, echo=FALSE, message=FALSE, warning=FALSE}
levels(imp$Item)

```

**The summary statistics of the number of tonnes of fruit imported by Ghana** 
```{r sum, echo=FALSE}
summary(imp$Value)
```
The results above reveals the following insights.<br>
**Max value**<br>
The highest quantity of fruits Ghana imported over 1998 to 2019 was **13,778 tonnes.**  
<br>
**Mean value**<br>
Over the 20 years (1998-2019), an **average of 384.4 tonnes of fruit** was imported into Ghana.

<br>
The following table shows the years in which Ghana's highest fruit imports occurred.




```{r 5, echo=FALSE, message=FALSE, warning=FALSE}
try_b=tibble(Item=c("Tomatoes","Apples","Pineapples","Grapes","Banana"),
             Year=c(2007,2013,2014,2013,2007),
             Value=c(13778,9656,472,403,327 ),
             Unit=c("tonnes","tonnes","tonnes","tonnes","tonnes"),
             Partner.Countries=c("Burkina Faso", "South Africa","Cote d'Ivoire",
                                 "South Africa","Cote d'Ivoire"))
#DT::datatable(dat2%>%
 # select(Item,Year,Value,Unit)%>%
 #filter(dense_rank(desc(Value))<=5))

#Item,Value,Year
try_y=imp%>%select(Item,Year,Value,Unit,Partner.Countries)%>%
            group_by(Year)%>%
            #slice_max(Item)
            filter(dense_rank(desc(Item))<=5)

#reactable(try_y,
          #defaultSorted = "Value",
         # defaultSortOrder = "desc",
          #filterable = T,
          #defaultPageSize = 20)


          

```

```{r 2,echo=FALSE, message=FALSE, warning=FALSE}
library(DT)
datatable(try_b)
#summary(dat5$cat_Value)
```

```{r 3, message=FALSE, warning=FALSE, include=FALSE}
reactable::reactable(f)
```

```{r try, eval=FALSE, include=FALSE}
try_select=function(x){
  only_top5%>%
  filter(Partner.Countries==x)%>%
  ggplot(aes(x=Item,y=Per_item, fill=Item))+geom_col()+
scale_fill_brewer(palette="Set2")+
geom_label(aes(label=Per_item),color="white",fill="black")+
labs(title="The 5 Most Imported fruits from  to Ghana",
subtitle="Values represent Frequency")
}

try_select("South Africa") 
try_select("France")
try_select("Netherlands")
try_select("Belgium")
try_select("Egypt")

```
